---
title: "Data 607 Assignment 4"
author: "Bridget Boakye"
date: "2022-09-18"
output: html_document
---

#Overview: This lab explores identifying patterns and regular expressions through various methods. 

1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}

#Load and check list of majors data from Github

theURL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors_list <-read.csv(file = theURL, header = TRUE, sep = ",")

head(majors_list)
```

```{r}

#use grep function to find Data or Statistics in majors column of data frame 

grep('DATA|STATISTICS', majors_list$Major, value = TRUE)

```


2. Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}

#Create list of vector of the data, ie. fruits 

fruits <- c("bell pepper", "bilberry", "blackberry","blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

fruits
```

```{r}

#Use paste and shQuote functions to construct strings with double quotes and produce output using concatenate 

cat(paste(shQuote(fruits, type="cmd"), collapse=", "))

```

3. Describe, in words, what these expressions will match:

a. (.)\1\1
b. “(.)(.)\\2\\1”
c. (..)\1
d. “(.).\\1.\\1”
e. “(.)(.)(.).*\\3\\2\\1”

*Answer*:

a. (.)\1\1: matches character appearing 3 times in a row (ex. "yyy")
b. "(.)(.)\\2\\1": matches a pair of characters followed by another pair of characters in reversed order (ex. "cssc")
c. (..)\1: matches two repeated characters (ex. "jaja")
d. "(.).\\1.\\1": matches a character followed by any character, then the original character, any other character, then the original character (ex. "a2aua")
e. "(.)(.)(.).*\\3\\2\\1": matches three characters followed by 0 or more characters then the same three characters but in reverse order. (ex. "bcddcb") 

4. Construct regular expressions to match words that:

a. Start and end with the same character
b. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
c. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

*Answer*:

a. str_subset(words, "^(.)((.*\\1$)|\\1?$)")
b. str_subset("church", "([A-Za-z][A-Za-z]).*\\1")
c. str_subset(words, "([A-Za-z][A-Za-z]).*\\1")


